Corotating Dust Clumps Around Adolescent M-Dwarfs: Four Years of TESS

### 2023-06-27 paper/ms.tex
update ms

		\vspace{-0.45cm}
		\subfloat{
			\includegraphics[width=0.8\textwidth]{f1b.pdf}
		}
	\end{center}
		\vspace{-0.2cm}
	\caption{
		{\bf Complex quasiperiodic variables (CQVs)}:
		{\it Top:} Phase-folded TESS light curves of three CQVs.  Each is
		stacked over one month.  Gray are raw 2-minute data; black bins to
		300 points per cycle.  Periods in hours are in the bottom right of
		each panel.  In order left-to-right, the objects are LP 12-502
		(TIC 402980664; Sector~19), TIC 94088626 (Sector 10), and TIC
		425933644 (Sector~28).
		{\it Bottom:} Plausible cartoon models for the phenomenon.  The
		dust clump scenario seems most plausible, given
		the stability of the dips, the lack of observed infrared
		excesses, and the challenge of producing broadband opacity variations
		with only ionized hydrogen in the prominence scenario.
	}
	\label{fig:f1}
\end{figure*}

@@

\subsubsection{LP 12-502 observations}
While many CQVs had multiple sectors of adjacent or nearly-adjacent data
LP 12-502 (TIC 402980664; $d$=21\,pc, $J$=9.4,
$T$=11.1) stood out as an outlier among the CQV class.
Figure~\ref{fig:lp} shows all available data from Sectors
18, 19, 25, 26, 53, 58, and 59; the star was observed at 120-second
cadence whenever it was observable by TESS.
Measuring the PDM peak period over each sector independently yielded
an average $\langle P \rangle = 18.5560$\,hr, with a difference
between the maximum and minimum sector-level period of about one minute.
Based on this range, the period is therefore stable to at least one
minute ($\pm$0.017\,hr) over the three year baseline.
However, in detail, a period shift of $\pm$1\,minute yields major
phase drifts over the 3 year baseline; that time interval
corresponds to roughly 1/1000$^{\rm th}$ of a period; we have
observed 1500 cycles.
The achievable period precision, $\sigma_{\rm P}$, can be estimated as
\begin{equation}
  \sigma_{\rm P} = \frac { \sigma_{\rm \phi} P } { N_{\rm baseline} },
\end{equation}
for $N_{\rm baseline}$ the number of cycles in the observed baseline and
$\sigma_{\rm phi}$ the phase precision with which any one feature (e.g.~a
dip) can be tracked.
Assuming $\sigma_{\rm phi} \approx 0.01$ yields an expected period precision
$\sigma_{\rm P} \approx 0.45 {\rm sec} \approx 1.2\times 10^{-4}\,{\rm hr}$.
By visually fine-tuning over a grid in period, we eventually 
found $P=18.5611 \pm 0.0001$\,hr, which seems to track certain
features in the LP 12-502 light curve over its entire baseline.

@@

\begin{itemize}
	%
	\item CQVs are {\it quasi}periodic.  While the mean periods in our
	sample seem to remain fixed over the span of available observations, the light curve
	shapes themselves evolve, similar to rotating stars.
	%
	\item The same CQV can show dips with closely-spaced periods.  
	LP 12-502, for instance, showed dips with four distinct periods within $\pm 0.3\%$
	of the fundamental period, sometimes simultaneous, and each lasting for up to 50 cycles.
	%
	\item The CQV phenomenon persists for at least 150\,Myr, based on the
	existence of CQVs in AB Dor, the Pleiades, and Psc-Eri.  It may extend to
	even older ages, however the lack of detected CQVs in the Hyades and
	Praesepe suggest that the phenomenon is unique to the pre-main-sequence
	for M-dwarfs.
	%
	\item CQVs evolve over timescales that are both secular ($>$100\,cycles) and impulsive ($<$1 cycle).
	``State-switches'' can cause dips to collapse instantaneously, and are often (but not exclusively)
	linked with observed optical flares.
	Dip growth however seems to happen over durations of at least ten cycles.
	%
	\item The population-averaged duty cycle for CQVs seems to be $\approx 75$\%, based on 
	the fraction of bonafide CQVs that ``turned off'' during TESS re-observations.
\end{itemize}

>>>>

		\vspace{-0.45cm}
		\subfloat{
			\includegraphics[width=0.8\textwidth]{f1b.pdf}
		}
	\end{center}
		\vspace{-0.3cm}
	\caption{
		{\bf Complex quasiperiodic variables (CQVs)}:
		{\it Top:} Phase-folded TESS light curves of three CQVs.  Each is
		stacked over one month.  Gray are raw 2-minute data; black bins to
		300 points per cycle.  Periods in hours are in the bottom right of
		each panel.  In order left-to-right, the objects are LP 12-502
		(TIC 402980664; Sector~19), TIC 94088626 (Sector 10), and TIC
		425933644 (Sector~28).
		{\it Bottom:} Plausible cartoon models for the phenomenon.  The
		dust clump scenario seems most plausible, given
		the stability of the dips, the lack of observed infrared
		excesses, and the challenge of producing broadband opacity variations
		with only ionized hydrogen in the prominence scenario.
	}
	\label{fig:f1}
\end{figure*}

@@

\subsubsection{LP 12-502 observations}
While many CQVs had multiple sectors of adjacent or nearly-adjacent data
LP 12-502 (TIC 402980664; $d$=21\,pc, $J$=9.4,
$T$=11.1) stood out as an outlier among the CQV class.
Figure~\ref{fig:lp} shows all available data from Sectors
18, 19, 25, 26, 53, 58, and 59; the star was observed at 120-second
cadence whenever it was observable by TESS.
Measuring the PDM peak period over each sector independently yielded
an average $\langle P \rangle = 18.5560$\,hr, with a difference
between the maximum and minimum sector-level period of about one minute.
Based on this range, the period is therefore stable to at least one
minute ($\pm$0.017\,hr) over the three year baseline.
However, in detail, a period shift of $\pm$1\,minute yields major
phase drifts over the 3 year baseline; that time interval
corresponds to roughly 1/1000$^{\rm th}$ of a period; we have
observed 1500 cycles.
The achievable period precision, $\sigma_{\rm P}$, can be estimated as
\begin{equation}
  \sigma_{\rm P} = \frac { \sigma_{\rm \phi} P } { N_{\rm baseline} },
\end{equation}
for $N_{\rm baseline}$ the number of cycles in the observed baseline and
$\sigma_{\phi}$ the phase precision with which any one feature (e.g.~a
dip) can be tracked.
Assuming $\sigma_{\rm phi} \approx 0.01$ yields an expected period precision
$\sigma_{\rm P} \approx 0.45 {\rm sec} \approx 1.2\times 10^{-4}\,{\rm hr}$.
By visually fine-tuning over a grid in period, we eventually 
found $P=18.5611 \pm 0.0001$\,hr, which seems to track certain
features in the LP 12-502 light curve over its entire baseline.

@@

\begin{itemize}
	%
	\item CQVs are {\it quasi}periodic.  While the mean periods in our
	sample seem to remain fixed over the span of available observations, the light curve
	shapes themselves evolve, similar to rotating stars.
	%
	\item The same CQV can show dips with very similar but clearly distinct periods.  
	LP 12-502, for instance, showed dips with four distinct periods within $\pm 0.3\%$
	of the fundamental period, sometimes simultaneous, and each lasting for up to 50 cycles.
	%
	\item The CQV phenomenon persists for at least 150\,Myr, based on the
	existence of CQVs in AB Dor, the Pleiades, and Psc-Eri.  It may extend to
	even older ages, however the lack of detected CQVs in the Hyades and
	Praesepe suggest that the phenomenon is unique to the pre-main-sequence
	for M-dwarfs.
	%
	\item CQVs evolve over timescales that are both secular ($>$100\,cycles) and impulsive ($<$1 cycle).
	``State-switches'' can cause dips to collapse instantaneously, and are often (but not exclusively)
	linked with observed optical flares.
	Dip growth however seems to happen over durations of at least ten cycles.
	%
	\item The population-averaged duty cycle for CQVs seems to be $\approx 75$\%, based on 
	the fraction of bonafide CQVs that ``turned off'' during TESS re-observations.
\end{itemize}
